---
title: "Introduction to JSON for the DNAnexus Platform"
author: "Ted Laderas"
institute: "DNAnexus"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
  powerpoint_presentation:
    slide_level: 1
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

prev_path <- Sys.getenv("PATH")

Sys.setenv(PATH = paste(prev_path, "/opt/homebrew/bin/", sep = ":"))
```

```{r xaringan-themer, include = FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "midnightblue",
  header_font_google = google_font("Josefin Sans"),
  link_color = "steelblue",
  text_font_size = "24px",
  footnote_font_size = "20px"
)
```

# Learning Objectives

1. **Explain** how JSON is used on the DNAnexus platform
2. **Articulate** the basic elements of JSON
3. **Describe** and **read** basic JSON structures that are used on the platform
4. **Parse** JSON responses from the platform using `jq` and pipes

---
# Refresher about DNAnexus architecture

- Your computer
- API Server
- Worker

---
# What is JSON?

- **J**ava**S**cript **O**bject **N**otation
- Common format for communicating with Application Program Interface (API) servers and receiving results/data from that API
- Used to access DNAnexus API servers

---
# Why is knowing about JSON useful?

- Reading and modifying JSON is at the heart of building and running apps
- Understanding JSON responses from the API will help you debug jobs
- Batch submissions: running the same app on multiple files
    - Find which jobs have failed and why
    - Run the failed jobs again
    
---
# Whether we use the UI or the command line, we are communicating with the platform using JSON

---
# Components of a JSON file

- Containers: 
  - `{}`
- key:value pairs
  - `"$dnanexus_link": "file-G4x7GX80VBzQy64k4jzgjqgY"`

.footnote[http://json-schema.org/draft-04/json-schema-core.html]

---
# key/value pairs

- have the format: `"key": "value"`
- example:

```
"dnanexus_link": "file-G4x7GXQ0VBzZxFxz4fqV120B"
```
---
# key/container

```
"report_html": {
    "dnanexus_link": "file-G4x7GX80VBzQy64k4jzgjqgY"
  }
```


---
# Anatomy of a JSON document

```
{
  "report_html": {
    "dnanexus_link": "file-G4x7GX80VBzQy64k4jzgjqgY"
  },
  "stats_txt": {
    "dnanexus_link": "file-G4x7GXQ0VBzZxFxz4fqV120B"
  }
}
```

---
# Building using JSON
  
- Building Apps
  - `dxapp.json` 
  - customize your app options
- Building Workflows
  - `dxworkflow.json`
  - Customizing instances for each step
  - Specifying files/software

---
# How is JSON used on the DNAnexus Platform?

- API requests
  - `dx find data`
  - `dx run` 
- API responses
  - Job Status (`dx watch`)
  - Job Results 

---
# Some advice

- Don't write JSON by hand from scratch
- Use an Editor!
    - Visual Studio Code
- Use JSON checking tools to make sure your JSON is well formed
  
---
# What is `jq`?

- Command line utility for working with JSON files
- Main functionality: filter
- Lets you extract and manipulate elements from a json file

---
# Install `jq` on your machine

- Need to have Anaconda installed first

```
conda install -c conda-forge jq
```

---
# filters - get first level

```
jq ".report_html" example.json
```

```
{
  "report_html": {
    "dnanexus_link": "file-G4x7GX80VBzQy64k4jzgjqgY"
  }
}
```

```
{
"dnanexus_link": "file-G4x7GX80VBzQy64k4jzgjqgY"
}
```

---
# Try this code out - what will it extract from example.json?

.pull-left[
```
jq '.stats_txt' data/example.json
```
]

.pull-right[
```
{
  "report_html": {
    "dnanexus_link": "file-G4x7GX80VBzQy64k4jzgjqgY"
  },
  "stats_txt": {
    "dnanexus_link": "file-G4x7GXQ0VBzZxFxz4fqV120B"
  }
}
```
]
---
# Actual Output

```{bash}
jq '.stats_txt' data/example.json
```

---
# Going down a level

.pull-left[
```
jq '.stats_txt.dnanexus_link' data/example.json
```
]

.pull-right[
```
{
  "stats_txt": {
    "dnanexus_link": "file-G4x7GXQ0VBzZxFxz4fqV120B"
  }
}
```
]

---
# Actual output

```{bash}
jq '.stats_txt.dnanexus_link' data/example.json
```


---
# Format JSON: `jq .`

- First step to helping you read JSON
- Usually responses are sent as a single line

```{bash}
echo '{"title":"This is very", "description": "hard to read" }' | jq "." 
```

---
# `. | keys`

- JSON files can be nested
- Let's look at the first level of keys

```{bash}
jq '. | keys' data/fastqc.json 
```

---
# JSON as input

```
dx run --json-input job_inputs.json
```

---
# Extract the categories from `dx describe`

```{bash}
jq '.categories' data/fastqc.json
```

---
# Things to Try

- Get the name of all of the apps run in a project:

```{bash}
dx find executions --json | jq ".[] | .name"
```
---
# Get the citation for an app

```{bash}
jq '.details.citations' data/fastqc.json
```

---
# Find all the first-level keys

```{bash}
jq '. | keys' data/rap-jobs.json
```

---

```{bash}
jq ".[] | keys" data/rap-jobs.json
```

---
# Cycle through the app names

```{bash}
jq '.name' data/jobs.json  
```

---
# Find the state of all jobs

```{bash}
jq ".[].state" data/rap-jobs.json
```

---
# Doing something with all of the jobs

```
while IFS= read -r obj; do
    ...
done < <(jq -c '.[]' data/rap-jobs.json)
```

---
# Piping 

- Pipe the output of `dx describe` into `jq`
- Find all outputs set for a job

```
dx describe job-G4x7G5j0B3K2FKzgP654ZqpK --json | jq -r .output
```

---
# Piping `dx describe` into `jq`

- Find all inputs set for a job, including default values set by the application

```{bash}
dx describe job-G4x7G5j0B3K2FKzgP654ZqpK --json | jq -r .input
```


---
# Find all inputs for an app

```{bash}
jq '.inputSpec' data/dxapp.json
```

---
# Find all generated outputs for a job

```{bash}
jq '.output' data/job.json
```

---
class: center, middle
# JSON is the lingua franca of the Internet
## Master it, and you open the Internet up
